karate

### 2021-11-06 inst/karate.Rmd
use sticker in Rmd
work on Zermelo's iterative scheme for pairwise comparisons

Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).

@@

```{r converttolikelihood}
karate <- hyper2()
for(i in seq_len(nrow(karate_table))){

@@

  karate[p1] %<>% inc(w1)
  karate[p2] %<>% inc(w2)
  karate[c(p1,p2)] %<>% dec(w1+w2)
}
karate
karate_maxp <- maxp(karate)
karate_maxp
```

>>>>

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).
Short story: the dataset has too many degrees of freedom to be easily
analysed with `hyper2` and Zermelo's iteration scheme as implemented
by the `BradleyTerry` package exploits the special pairwise structure
of the likelihood function very effectively.

@@

```{r converttolikelihood,cache=TRUE}
karate <- hyper2()
for(i in seq_len(nrow(karate_table))){

@@

  karate[p1] %<>% inc(w1)
  karate[p2] %<>% inc(w2)
  karate[c(p1,p2)] %<>% dec(w1+w2)
}
summary(karate)
```

In the summary above (in the summary of bracket lengths), we see how
the likelihood function comprises only singletons [which appear on the
numerator] and pairs [on the denominator].

```{r}
karate_maxp <- maxp(karate)
```

```{r}
dotchart(log10(karate_maxp))
```

# MM analysis

We will define a matrix of win-loss statistics:

```{r}
pk <- pnames(karate)
M <- matrix(0,length(pk),length(pk))
rownames(M) <- pk
colnames(M) <- pk
o <- karate_table # saves typing
for(i in seq_len(nrow(o))){
  M[o[i,1],o[i,3]] %<>% `+`(o[i,2])
  M[o[i,3],o[i,1]] %<>% `+`(o[i,4])
}
```

```{r,cache=TRUE}
pM <- pairwise(M)
pM == karate  # consistency check; should be TRUE
```

Now use Zermelo's iteration method and compare with `hyper2`'s `maxp()`:

```{r}
iM <- zermelo(M)
loglik(iM,karate)
loglik(karate_maxp,karate)
```

Above we see that Zermelo gives a much more likely point in parameter
space than `maxp()`.

```{r}
par(pty='s')
plot(log10(iM),log10(karate_maxp),asp=1,xlim=c(-4,0),ylim=c(-4,0))
abline(0,1)
grid()
```
